Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-06-26 AM+draft.tex
27June18 methods v2

BigRR, LMM-Lasso, and glmnet  implement different regression-based regularisation methods. 
Regularisation methods can handle parameter estimation problems where the number of predictors is far greater than the number of 
samples. When applied to association mapping, they allow the effect of all snp to be modelled simultaneously. 
BigRR implements generalized ridge regression.
In ridge regression, the coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridge regression, the amount of shrinkage is allowed to vary. LMM-Lasso 
implements lasso within a linear mixed model 
framework. Unlike in ridge regression, lasso performs parameter estimation and variable selection, simultaneously.   The 
coefficients of "unimportant" predictors are shrunk to exactly zero.
Glmnet is based on elastic net, a regularisation method that is an improvement over ridge regression and lasso (ref needed). 
All three methods are sophisticated pieces of statistical machinery. The challenge is that while they have the capacity to estimate the 
effect sizes of snp across an entire genome simultaneously, obtaining the statistical significance of these snp is problematic. 
Permutation has been suggested as a way of calculating significance 
empirically \cite{shen2013novel} but we instead opted for stability selection (see below).

Regularisation methods have dominated 
the development of multiple locus approaches.  They allow the effect of all the snp to be modelled simultaneously. 
By introducing a penalty function that balances bias and variance, regularisation methods can 
handle parameter estimation problems where the number of predictors is far greater than the number of samples.  
One of the challenges in using BigRR for association mapping is that while it has the capacity to estimate the effect sizes of all the snp simultaneously, 
obtaining their statistical significance is difficult. 
Permutation has been suggested as a way of calculating significance 
empirically \cite{shen2013novel} but we instead opted for stability selection.

MLMM, of the five multiple locus methods, is the closest in philosophy to Eagle. It too is based on a model building philosophy 
and builds the best linear mixed model via a forward selection strategy. However, MLMM does not make use of dimension 
reduction which adds to its computational cost. Also, its model building process is different. Eagle 
identifies the $(s+1)$th snp that is in strongest association with the trait by maximising a score statistic. MLMM 
identifies this snp by constructing a separate linear mixed model for each unidentified snp and calculating its statistical significance in
the fixed effects part of the model. The snp with the highest significance is included in the next iteration of the model building process.

R2VIM is quite different to the other association mapping methods. It implements random forests for association mapping, 
a machine learning algorithm to measure the 
importance of the snp-trait associations. 
There is no relationship between relative importance scores and significance levels. Consequently, 
setting an appropriate genome-wide threshold is problematic. We avoid this problem by focusing, in the simulation study, 
only on the relationship between threshold level and its false discovery rate and power.

>>>>

BigRR, LMM-Lasso, and glmnet  implement three different regression-based regularisation methods; generalized ridge regression, 
lasso, and elastic net, respectively. 
Regularisation methods can handle parameter estimation problems where the number of predictors is far greater than the number of 
samples. This is particularly advantageous in the context of association mapping as it allows the effects of snp across an entire 
genome to be estimated simultaneously. 
All three methods are sophisticated pieces of statistical machinery. The limitation though is that the significance of an effect, which 
is needed in association mapping,  can not be calculated easily. 
 A possible solution is to estimate the significance of the effects via permutation  \cite{shen2013novel}  
but we have instead opted for stability selection (see below).

MLMM, of the five multiple locus methods, is the closest in philosophy to Eagle. It too is based on a model building philosophy 
and builds the best linear mixed model via a forward selection strategy. However, MLMM does not make use of dimension 
reduction which adds to its computational cost. Also, its model building process differs from that of Eagle's. Eagle 
identifies the $(s+1)$th snp that is in strongest association with the trait by finding the snp effect with the maximum score statistic. 
MLMM instead identifies the snp, that when included as a fixed effect in the model, has the highest statistical 
significance. This neccesitates the construction and fitting of a separate linear mixed model for each unidentified snp. MLMM does this efficiently.

R2VIM implements random forests and a new way of measuring the importance of a snp.  In random forests, the worth of a predictor is measured, empirically, by calculating its importance score. It is from these importance scores that snp can be ordered in terms of their strength of association with a trait. The challenge though is in knowing what proportion of the 
highest ordered snp are in true association. R2VIM addresses this, partially, by calculating a relative importance score for a snp. However, 
there is no relationship between relative importance scores and significance levels, making the setting of a threshold difficult.

======================= The above needs a bit more detail may need to look up papers ================
